On Verification Patterns for Low-Level Systems via Modal Abstractions

### 2025-09-11 paper.tex
+

\textit{Contributions} We argue that modal abstractions can be used to identify and abstract system verification challenges. We justify our perspective by discussing prior systems that have successfully used modalities for system verification, arguing that they fit into the verification design pattern we articulate, and explaining how this approach might apply to other systems' verification challenges.
\begin{itemize}
\item \textbf{Identifying System Verification Challenges}: We start with identifying common patterns in system verification: \emph{virtualization, sharing, and translation}.
\item \textbf{Introducing Modal \emph{Resource Contexts}}: Then we discuss the concept of \emph{resource} which has already been an essential concept in the design of systems. Inspired by the concept of resource in the systems, we define what a resource and its context are in the modal abstractions.
\item \textbf{Introducing \emph{Nominals for System Resources}}:
Existing logics for systems software must frequently identify
individual resources within a given context ---
nominalization (in the sense of hybrid logic) enables identifying a resource in its context. 
For example, a filesystem transaction is a context of resources for in-memory updated disk blocks
which have not yet been saved to disk.
The transaction identifier is used to associate a transaction with a disk-block to be persisted so that, 
in case of a crash while persisting updated disk-blocks, the filesystem can rollback the 
\emph{already persisted} disk-blocks of the transaction, and reach to the previous consistent disk state. 
We discuss other examples to contextualize the existing systems surveyed.
% to be able to do so, both the updated in-memory disk blocks and the transaction must refer to the transaction 
% identifier -- \emph{strong nominalization}. 
% in another example, virtual memory references (resources) in 
% an address space (resource context), which can be uniquely identified with a root address (the nominal) 
% of its page-table tree, are \emph{agnostic} to the address space that they are in. 
% However, they can only be accessed (be valid) in the address space to which they are agnostic. 
% However, then an address space switch happens, the virtual memory references of the previous address 
% space must be made inaccessible. To be able to do so, although virtual memory references do not hold any 
% piece of information related to their address space, we still have to associate them. 
% We call this kind of unilateral nominalization -- \emph{weak nominalization}.
% \todo[inline,color=cyan]{I left this paragraph alone, mostly, but it's too detailed for the new intro section position here}
\item \textbf{Taxonomy of Current Modal Approaches in System Verification}: Based on these concepts defined, we summarize contemporary verification efforts using modal abstractions. We choose them from different domains, for example, reasoning about weak memory models and storage persistence ~\cite{tejthesis,perennialgit,tejperennial19,larsnextgen25,fsl,fsl++,derekrustbelt20,kuru2025modalabstractionsvirtualizingmemory,amalreal2024} because we would like to justify that our \emph{definitions} are not domain dependent.
\end{itemize}

>>>>

\textit{Contributions} We argue that modal abstractions can be used to identify and abstract system verification challenges. We justify our perspective by discussing prior systems that have successfully used modalities for system verification, arguing that they fit into the verification design pattern we articulate, and explaining how this approach might apply to other systems' verification challenges.
\begin{itemize}
\item \textbf{Identifying System Verification Challenges}: We start with identifying common patterns in system verification: \emph{virtualization, sharing, and translation}.
\item \textbf{Introducing Modal \emph{Resource Contexts}}: Then we discuss the concept of \emph{resource} which has already been an essential concept in the design of systems. Inspired by the concept of resource in systems, we define what a resource and its context are in the modal abstractions. For example, the \emph{transaction} abstraction, which behaves like a container for in-memory dirty disk pages, can be considered as a resrouce context for any dirty in-memory disk page resource associated with it.
\item \textbf{Introducing \emph{Nominals for System Resources}}:
Existing logics for systems software must frequently identify
individual resources within a given context ---
nominalization (in the sense of hybrid logic) enables identifying a resource in its context. 
For example, a filesystem transaction is a context of resources for in-memory updated disk blocks
which have not yet been saved to disk.
The transaction identifier is used to associate a transaction with a disk-block to be persisted so that, 
in case of a crash while persisting updated disk-blocks, the filesystem can rollback the 
\emph{already persisted} disk-blocks of the transaction, and reach to the previous consistent disk state. 
We discuss other examples to contextualize the existing systems surveyed.
% to be able to do so, both the updated in-memory disk blocks and the transaction must refer to the transaction 
% identifier -- \emph{strong nominalization}. 
% in another example, virtual memory references (resources) in 
% an address space (resource context), which can be uniquely identified with a root address (the nominal) 
% of its page-table tree, are \emph{agnostic} to the address space that they are in. 
% However, they can only be accessed (be valid) in the address space to which they are agnostic. 
% However, then an address space switch happens, the virtual memory references of the previous address 
% space must be made inaccessible. To be able to do so, although virtual memory references do not hold any 
% piece of information related to their address space, we still have to associate them. 
% We call this kind of unilateral nominalization -- \emph{weak nominalization}.
% \todo[inline,color=cyan]{I left this paragraph alone, mostly, but it's too detailed for the new intro section position here}
\item \textbf{Taxonomy of Current Modal Approaches in System Verification}: Based on these concepts defined, we summarize contemporary verification efforts using modal abstractions. We choose them from different domains, for example, reasoning about weak memory models and storage persistence ~\cite{tejthesis,perennialgit,tejperennial19,larsnextgen25,fsl,fsl++,derekrustbelt20,kuru2025modalabstractionsvirtualizingmemory,amalreal2024} because we would like to justify that our \emph{definitions} are not domain dependent.
\end{itemize}
